---
title: "Untitled"
author: "Moiya Josephs"
date: "5/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(recommenderlab)
library(tidyverse)
library(reshape2)
```

# Overview

```{r}
data("MovieLense")
movielense <- MovieLense
remove(MovieLense)
```


Preview what the data looks like: 

```{r}
getRatingMatrix(movielense[1:10,1:4])
```

```{r}
summary(getRatings(movielense))

data.frame(ratings = getRatings(movielense)) %>%
  ggplot(aes(ratings)) + geom_bar(width = 0.75)
```

```{r}
train_proportion <- .75
min(rowCounts(movielense))
```
```{r}
items_per_test_user_keep <- 10
# What's a good rating for a binary split?
good_threshold <- 4
```

```{r}
# Building a Recommender System with R by Gorakala and Usuelli. Ch.4 pp 77 - 83
set.seed(12345)
model_train_scheme <- movielense %>%
  evaluationScheme(method = 'split', # single train/test split
                   train = train_proportion, # proportion of rows to train.
                   given = items_per_test_user_keep, # shouldn't keep n rec. items > min(rowCounts(movie_r))
                   goodRating = good_threshold, # for binary classifier analysis.
                   k = 1)
```


```{r}
# Building a Recommender System with R by Gorakala and Usuelli. Ch.4 pp 84
model_params <- list(method = "cosine",
                     nn = 10, # find each user's 10 most similar users.
                     sample = FALSE, # already did this.
                     normalize = "center")

model1 <- getData(model_train_scheme, "train") %>% #only fit on the 75% training data.
  Recommender(method = "UBCF", parameter = model_params)
```

```{r}
# 5.5 - 5.6. Evaluation of predicted ratings in recommenderLab vignette. can use n = for predicting TopN or type = for predicting ratings.
# https://cran.r-project.org/web/packages/recommenderlab/vignettes/recommenderlab.pdf
model1_pred <- predict(model1, getData(model_train_scheme, "known"), type = "ratings")
model1_pred
```

```{r}
test_error <- calcPredictionAccuracy(model1_pred, getData(model_train_scheme, "unknown"), byUser = TRUE)
head(test_error)
```
```{r}
eval_movies <- evaluationScheme(data = movielense, 
                      method = "cross-validation", 
                      k = 10,
                      given = 15, 
                      goodRating = 4)
eval_movies
```
```{r}
train_movies <- getData(eval_movies, "train")
known_movies <- getData(eval_movies, "known")
unknown_movies <- getData(eval_movies, "unknown")
```

```{r}

data.frame(ratings = getRatings(train_movies)) %>%
  ggplot(aes(ratings)) + geom_bar(width = 0.75)
```
```{r}

data.frame(ratings = getRatings(known_movies)) %>%
  ggplot(aes(ratings)) + geom_bar(width = 0.75)
```
```{r}

data.frame(ratings = getRatings(unknown_movies)) %>%
  ggplot(aes(ratings)) + geom_bar(width = 0.75)
```


```{r}
ibcf <- 
  train_movies %>%
  Recommender(method = "IBCF") 

ibcf_eval <- ibcf %>% 
  predict(known_movies, type = "ratings") %>% 
  calcPredictionAccuracy(unknown_movies)
print(ibcf_eval)
```

```{r}
ubcf <- 
  train_movies %>%
  Recommender(method = "UBCF")

ubcf_eval <- ubcf %>% 
  predict(known_movies, type = "ratings") %>% 
  calcPredictionAccuracy(unknown_movies)

print(ubcf_eval)

```
```{r}
pop <- 
  train_movies %>%
  Recommender(method = "POPULAR")

pop_eval <- pop %>% 
  predict(known_movies, type = "ratings") %>% 
  calcPredictionAccuracy(unknown_movies)

print(pop_eval)

```

```{r}
random <- 
  train_movies %>%
  Recommender(method = "RANDOM")

random_eval <- random %>% 
  predict(known_movies, type = "ratings") %>% 
  calcPredictionAccuracy(unknown_movies)
print(random_eval)

```

```{r}
svd <- 
  train_movies %>%
  Recommender(method = "SVD")

svd_eval <- svd %>% 
  predict(known_movies, type = "ratings") %>% 
  calcPredictionAccuracy(unknown_movies)
print(svd_eval)


```

# Testing the best predictor


# Analyze the data



# Clean the data so that it looks like movie lense

The data gives information that will just confuse the recommendation algorithm.

Slice out all the important data.





```{r}
movie1<- read.delim("C:/Users/josep/movie data/2.txt", header = FALSE, sep = ",", dec = ".")
# add movie id
movie1<- movie1[,c(1,2)]
movie_id <- rep(c(1),547)
movie1<- cbind(movie_id,movie1)
```

```{r}
movie_id <- rep(c(2),145)
movie2<- read.delim("C:/Users/josep/movie data/3.txt", header = FALSE, sep = ",", dec = ".")
movie2<- movie2[,1:2]
movie2<- cbind(movie_id,movie2)
```


```{r}
movie_id <- rep(c(3),2012)
movie3<- read.delim("C:/Users/josep/movie data/4.txt", header = FALSE, sep = ",", dec = ".")
movie3<- movie3[,1:2]
movie3<- cbind(movie_id,movie3)
```


```{r}
movies<- rbind(movie1, movie2, movie3)

```

```{r}
dim(movies)

```


```{r}
ratingsMatrix <- dcast(movies, V1~movie_id, value.var = "V2" )

```

```{r}
ratingsMatrix <- as.matrix(ratingsMatrix[-c(1)])
```
  
  
```{r}
# Convert rating matrix into a recommenderlab sparse matrix
ratingsMatrix <- as(ratingsMatrix, "realRatingMatrix")

ratingsMatrix
```

```{r}
getRatingMatrix(ratingsMatrix)
```   


```{r}
vec_ratings <- as.vector(ratingsMatrix@data)

# Unique ratings
unique(vec_ratings)
```

```{r}
evaluation <- evaluationScheme(ratingsMatrix, method = "split", train = 0.8, given = 1)

evaluation
```

```{r}
train <- getData(evaluation, "train")
train
```

```{r}
test_known <- getData(evaluation, "known")
test_known
```

```{r}
test_unknown <- getData(evaluation, "unknown")
test_unknown
```

```{r}
set.seed(123)
rec_als <- Recommender(data = train, method = "IBCF")

pred_als_acr <- predict(object = rec_als, newdata = test_unknown, type = "ratings")
pred_als_acr
```

```{r}
model1_pred <- predict(rec_als, getData(evaluation, "known"), type = "ratings")
model1_pred
```



```{r}
getRatingMatrix(model1_pred)
```

```{r}

test_error <- calcPredictionAccuracy(model1_pred, test_unknown)
test_error
```

From research is is because the data is to sparce it will show NaN.



-----
Do it on Ratings Beuty

```{r}
ratings <- read.csv("C:/Users/josep/OneDrive/Documents/Data-607-Projects/Final Project/ratings_Beauty.csv", header = TRUE, sep = ",", dec = ".")
ratings<- sample_n(ratingsBeauty, 1000)
```

```{r}
class(ratings$UserId); class(ratings$ProductId); class(ratings$Rating); class(ratings$Timestamp)
```

```{r}
ratings$UserId<- as.factor(ratings$UserId); 
ratings$ProductId <- as.factor(ratings$ProductId)
```

```{r}
ratingsMatrix <- sparseMatrix(as.integer(ratings$UserId), as.integer(ratings$ProductId), x = ratings$Rating)
colnames(ratingsMatrix) <- levels(ratings$ProductId)
rownames(ratingsMatrix) <- levels(ratings$UserId)
amazon <- as(ratingsMatrix, "realRatingMatrix")
```



```{r}
length(unique(ratingsBeauty$UserId))
length(unique(ratingsBeauty$ProductId))
```
```{r}


implode <- function(x) {
   x<-str_replace_all(x,"\\D","") 
    as.numeric(x)
    
}
implode("c(1,2)")




```
```{r}
unlisting<- function(x){
  x<-str_c(x,collapse='')
  as.numeric(x)
}

unlisting(c("1","2"))
```



```{r}

ratingsBeauty$UserId<- lapply(ratingsBeauty$UserId, utf8ToInt)
ratingsBeauty$UserId<- lapply(ratingsBeauty$UserId, as.character)
```

```{r}
ratingsBeauty$UserId<- lapply(ratingsBeauty$UserId, unlisting)
```




```{r}
ratingsBeauty$ProductId<- lapply(ratingsBeauty$ProductId, utf8ToInt)
ratingsBeauty$ProductId<- lapply(ratingsBeauty$ProductId, as.character)
```

get the integers 

```{r}
ratingsBeauty$ProductId<- lapply(ratingsBeauty$ProductId, unlisting)
```


make sure the uniue values stay unique 


```{r}
length(unique(ratingsBeauty$UserId))
length(unique(ratingsBeauty$ProductId))
```
```{r}
class(ratingsBeauty$UserId); class(ratingsBeauty$ProductId); class(ratingsBeauty$Rating); class(ratingsBeauty$Timestamp)

```
```{r}
ratingsBeauty$UserId <- as.numeric(ratingsBeauty$UserId)

ratingsBeauty$ProductId <- as.numeric(ratingsBeauty$ProductId)
```

```{r}
sum(is.na(ratingsBeauty$UserId))
sum(is.na(ratingsBeauty$ProductId))
```
```{r}
class(df$UserId); class(df$ProductId); class(df$Rating); class(df$Timestamp)

```


```{r}
df$UserId <- as.factor(df$UserId)

df$ProductId <- as.factor(df$ProductId)
```


```{r}
df <- as.data.frame(as.matrix(amazon@data))
df
```


```{r}
df$UserId <- rownames(df)
df <- df %>% gather(key = ProductId, value = Rating, -UserId) %>% filter(Rating != 0)
df
```



```{r}
ratingsMatrix <- sparseMatrix(as.numeric(df$UserId), as.integer(df$ProductId), x = df$Rating)
colnames(ratingsMatrix) <- levels(df$ProductId)
rownames(ratingsMatrix) <- levels(df$UserId)
amazon <- as(ratingsMatrix, "realRatingMatrix")
```

  
```{r}
# Convert rating matrix into a recommenderlab sparse matrix
ratingsBeautyMatrix <- as(ratingsBeautyMatrix, "realRatingMatrix")

ratingsBeautyMatrix
```
```{r}

```


```{r}
getRatings(ratingsBeautyMatrix)
```

``

```{r}
beautyevaluation <- evaluationScheme(ratingsBeautyMatrix, method = "split", train = 0.8, given = 1, goodRating= 4)

beautyevaluation
```

```{r}
btrain <- getData(beautyevaluation, "train")
btrain
```

```{r}
btest_known <- getData(beautyevaluation, "known")
btest_known
```

```{r}
btest_unknown <- getData(beautyevaluation, "unknown")
btest_unknown
```

```{r}
set.seed(12345)
rec_beauty <- Recommender(data = btrain, method = "POPULAR")

pred_beauty <- predict(object = rec_beauty, newdata = btest_unknown, type = "ratings")
pred_beauty
```

```{r}
evaluation <- evaluationScheme(ratingsBeautyMatrix, method = "split", train = 0.8, given = 1)

model1_pred <- predict(rec_beauty, getData(evaluation, "known"), type = "ratings")
model1_pred
```



```{r}
getRatingMatrix(ratingsBeautyMatrix[1:200,1:4])
```

```{r}

test_error <- calcPredictionAccuracy(model1_pred, getData(evaluation, "unknown"))
test_error
```

